---
params:
  target_study: "ryegrass"
title: "Bio-Efficacy Dose-Response Report: `r stringr::str_to_title(params$target_study)`"
author: "Keith Post"
date: last-modified
format: 
  html:
    theme: cosmo
    toc: true
    code-fold: true
    self-contained: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

# Import libraries 
library(here)
library(tidyverse)



# 1. Load suitcase
my_suitcase <- here("output", "models", paste0(params$target_study, "_full_suitcase.rds")) %>%
                      readRDS()


# 2. Define a "Pass" flag
pass_structural <- my_suitcase$status != "Failed"


# 3. Only attempt to define downstream objects if validation passed
if(pass_structural) {
  target_study <- params$target_study
  fp_audit_plot <- paste0("output/plots/", target_study, "_audit_plot.png")
  fp_final_plot <- paste0("output/plots/", target_study, "_final_curve.png")

  # Only try to read the CSV if modeling actually ran
  if(!is.null(my_suitcase$modeling$results)) {
    df_stats <- here("output", "tables", paste0(target_study, "_summary_stats.csv")) %>%
      read.csv()
  }
  
  # Define display objects
  snr <- round(my_suitcase$snr %||% 0, 2)
  dose_ratio <- round(my_suitcase$dose_ratio %||% 0, 2)
  n_outliers <- length(my_suitcase$outlier_indices)
  at_plateau <- my_suitcase$at_plateau
  
  model_name <- my_suitcase$modeling$winner_name
  model_label <- my_suitcase$modeling$winner_label
  model_rationale <- my_suitcase$modeling$rationale
  model_table <- my_suitcase$modeling$table
}

```

<!-- "Fatal Error" Banner -->
`r if(my_suitcase$status == "Failed") {
"::: {.callout-important}\n## ❌ Critical Error: Pipeline Halted\n"
}`
`r if(my_suitcase$status == "Failed") {
  paste0("**Validation Message:** ", my_suitcase$message, "\n:::")
}`


<!-- SNR Banner -->
`r if(snr >= 5) {
"::: {.callout-note appearance='simple'}\n## [PASS] Robust Signal: This dataset provides high confidence for parameter estimation.\n:::"
} else if(snr >= 3) {
"::: {.callout-warning appearance='simple'}\n## [REVIEW] Acceptable Signal: Results are reliable, but observe confidence intervals.\n:::"
} else {
"::: {.callout-important appearance='simple'}\n## [CAUTION] Low Signal: Model results should be considered exploratory.\n:::"
}`



## Data Quality & Suitability Audit
### Trial Robustness
* **Signal-to-Noise Ratio (SNR):** The analysis of the dataset returned an SNR of `r snr`, which suggests `r if(snr>=5){
"robust treatment effect with minimal background interference."
} else if(snr >= 3 & snr < 5){
"clear dose-response signal; suitable for reliable parameter estimation."
} else if(snr >= 1.5 & snr < 3){
"a clear detectable signal, but high variance may increase $ED_{50}$ uncertainty."
} else{
"a signal indistinguishable from noise. Modeling results should be considered exploratory."
}`

* **Dose Coverage:** The `r target_study` dataset has a dose ratio of `r dose_ratio`. This value indicates `r if(dose_ratio >= 50){
"exceptional dose distribution across multiple orders of magnitude."
} else if(dose_ratio >= 10 & dose_ratio < 50){
"standard dose coverage for robust non-linear parameter estimation."
} else if(dose_ratio >= 5 & dose_ratio < 10){
"risk of 'fragmented' curves; the model may struggle to fix the slope accurately."
} else{
"insufficient range; the resulting model may be under-determined."
}`


### Data Integrity
* **Outlier Count:** `r if(n_outliers==0){
paste("No outliers were found in the", target_study,"dataset.")
} else if(n_outliers==1){
paste("One outlier was found in the", target_study, "dataset.")
} else{
paste(n_outliers, "outliers were found in the", target_study, "dataset.")
}`

* **Plateau Status:** `r if(at_plateau){
paste("The dose range successfully captured the terminal response plateau, providing a high degree of confidence in the upper-limit estimation.")
}else{
"Warning: The biological response did not reach a stable plateau within the tested dose range. Estimations of maximum effect and relative potency ($ED_{50}$) may involve mathematical extrapolation."
}`

### Visual Audit Summary
The following figure illustrates the distribution of raw data points (replicates) relative to the calculated dose-means. This visualization allows for a direct assessment of the 'Signal' (the trend of the blue line) against the 'Noise' (the vertical spread of the grey points and error bars).

`r if(n_outliers > 0){
"Any points highlighted with **red X's** represent identified outliers that were evaluated during the quality control sweep."}`




```{r} 
#| label: fig-audit
#| eval: !expr pass_structural
#| fig-cap: "Visual Audit: Distribution of raw replicates across dose levels."
#| fig-align: center
#| out-width: "7in"

knitr::include_graphics(fp_audit_plot)

```


## Dose-Response Analysis & Parameter Estimation
## Modeling Methodology
The data were evaluated against a hierarchy of candidate models to determine the most statistically appropriate and biologically relevant representation of the dose-response relationship.

### Selection Framework
The best-fit model was determined using the following tiered decision logic:

1. **Non-Linear Sigmoidal Evaluation**:
   * Initial attempts were made to fit **Brain-Cousens (Hormetic)**, **Log-Logistic (Standard)**, and **Log-Logistic (Fixed Baseline)** models.
   * If multiple models converged, a **parsimony check** was performed using the Akaike Information Criterion (AIC). A more complex model was only selected if it improved the fit by at least **2 AIC units** over its simpler counterpart.
   * The resulting candidate was then required to demonstrate a **statistically significant slope** ($p < 0.05$).

2. **Linear Fallback**:
   * If sigmoidal models failed to converge or show significance, a **Linear Trend** was evaluated. This model was selected only if it achieved a significant dose-dependent slope ($p < 0.05$).

3. **No Detectable Response (Null)**:
   * If all previous attempts failed to detect a significant relationship, the dataset was categorized as having **No Detectable Response**, utilizing an intercept-only model.


### Model Selection
`r if(model_name=="Null") { 
paste("Evaluation of the dataset indicates", paste0("**", model_label, "**"), ". ", model_rationale) 
} else{ 
paste("The biological response was characterized using a", paste0("**", model_label, "**"), "function. This selection was based on the following criteria:", model_rationale) 
}`


### The Final Curve

```{r} 
#| label: fig-final
#| eval: !expr pass_structural && !is.null(my_suitcase$modeling$plot)
#| fig-cap: !expr "paste('Final Dose-Response Curve: Fitted model for', my_suitcase$metadata$y_name)"
#| fig-align: center
#| out-width: "7in"

knitr::include_graphics(fp_final_plot)

```


## Model Performance & Estimates
`r if(model_name=="Null") {
"No potency parameters could be calculated."
} else{
paste("The table below summarizes the statistical output and potency estimates for the", target_study, "dataset.")
}`

```{r}
#| label: table-potency
#| eval: !expr model_name != "Null"

model_table

```





